Online reprocessing simulation for thorium-fueled molten salt breeder reactor

### 2017-07-07 msbr_reproc.tex
figure spacing too tight

The thorium-fueled \gls{MSBR} was developed in the early 1970s by \gls{ORNL} 
specifically to realize the promise of the thorium fuel cycle which allows the 
use of natural thorium instead of enriched uranium as the fertile element. 
Thorium breeds the fissile $^{233}U$ and avoids uranium enrichment 
\cite{robertson_conceptual_1971}. In the matter of nuclear fuel cycle, the 
thorium cycle produces a reduced quantity of plutonium and \glspl{MA} 
compared to the traditional uranium fuel cycle. Consequently, it may 
significantly increase proliferation resistance when the \gls{MSR} operates in the 
breeder regime. The \glspl{MSR} also could be employed as converter reactor for 
transmutation spent fuel from \gls{LWR} or others.

Nowadays, interest to \gls{MSR} coming back after long-term break because of 
its unique characteristics and features of this type of reactor, such as online 
reprocessing and refueling. For the development of \gls{MSR} conception special 
computational analysis methods and codes needed due to completely different 
physics of liquid-fueled nuclear reactor comparing with traditional, 
solid-fueled, reactors. Most of the contemporary nuclear reactor physics codes 
do not able to perform depletion calculations in the online reprocessing 
regime. J.J. Powers (\gls{ORNL}) has suggested a novel method to conducting 
depletion simulation for \gls{MSR} with taking into account the online 
reprocessing and refueling based on the deterministic computer code NEWT in 
SCALE \cite{powers_new_2013}. This approach was later used by Jeong et al. to 
find equilibrium state for \gls{MSBR} and validate it with \gls{MCNP}/CINDER90 
model \cite{jeong_equilibrium_2016}. For the development of \gls{MSBR} 
research, this paper presents the single-cell model developed using a 
continuous-energy the Serpent 2 Monte Carlo reactor physics calculation code 
that was employed to find equilibrium core state, and several calculation 
results including the depletion calculation of the single-cell unit.

@@

\begin{figure}[h] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \vspace{-0.6em}
        \includegraphics[width=0.90\linewidth]{zone_I_mesh.png}
        \caption{MSBR unit cell of Zone I geometry.}
        \label{fig:zoneI}
\end{figure}

@@

\begin{figure}[hb] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \vspace{-0.8em}
        \includegraphics[width=1.03\linewidth]{keff.png}
        \caption{Infinite multiplication factor during 4 years depletion 
        simulation.}
        \vspace{-0.8em}
        \label{fig:keff}
\end{figure}

@@

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \vspace{-0.8em}
        \includegraphics[width=1.03\linewidth]{fuel_composition.png}
        \caption{Normalized number density of major isotopes during 4 years 
        depletion.}
        \label{fig:compos}
        \vspace{-0.8em}
\end{figure}

The analysis of the fuel salt composition variation gives clearer information 
about equilibrium core state. Figure~\ref{fig:compos} shows the normalized 
number density of important isotopes that have strong influence on the core 
neutronics at the beginning of each depletion time interval. The number density 
of protactinium is very low (less than $10^{16}$1/cm$^3$) but some small amount 
of it produced during 3-days reprocessing step. It can be seen that the 
multiplication factor reaches the equilibrium state at about 950  days. 
Figure~\ref{fig:rates} represents the rates of materials in online reprocessing 
flows during 4 years of depletion calculations. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\captionsetup[table]{
%  labelsep = newline,
%  name = TABLE, justification=justified,
%  singlelinecheck=false,%%%%%%% a single line is centered by default
%  labelsep=colon,%%%%%%
%  skip = \medskipamount}
%\begin{table}[h!]
%\centering
%\begin{tabular}{p{0.25\linewidth} p{0.3\linewidth} p{0.3\linewidth}} \toprule
%      Isotope & First 400 days     & After 400 days          
%\\ \midrule
%$^{232}$Th  & 3.704$\times10^{12}$ & 3.086$\times10^{12}$
%\\ \midrule
%$^{233}$Pa  & 3.704$\times10^{12}$ & 3.086$\times10^{12}$
%\\ \midrule
%$^{233}$U  & $1.05134\pm0.00116$ & $1.00736$
%\\
%\bottomrule
%\end{tabular}
%  \caption{Materials flows rates during \gls{MSBR} online reprocessing 
%  [1/s*cm$^3$].}
%  \label{tab:rates}
%\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[hb] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \vspace{-0.8em}
        \includegraphics[width=1.03\linewidth]{rates_fuel.png}
        \caption{Materials flows rates during online reprocessing.}
        \label{fig:rates}
        \vspace{-0.8em}
\end{figure}
From the graph below we can see that to keep reactor critical higher $^{233}$U 
flow rate required for first 400 days of operation and after that much smaller 
uranium-233 flow rate from the protactinium decay tank needed. The $^{232}$Th 
rate slightly decreases over 4 years of operation due to other than $^{233}$U 
fissile materials accumulation.

\begin{figure}[hb] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \vspace{-0.8em}
        \includegraphics[width=1.03\linewidth]{rates_outflow.png}
        \caption{Materials flows rates for the protactinium decay tank during 
        \gls{MSBR} online reprocessing.}
        \vspace{-0.6em}
        \label{fig:outflow}
        \vspace{-0.8em}
\end{figure}
As shown in Figure~\ref{fig:outflow}, the tank for protactinium decay 
accumulates $^{233}$Pa approximately 200 days and fresh fuel ($^{233}$U) flow 
established also after 200 days. Uranium produced in the tank by protactinium 
decay is separated by circulation of the salt through a flourinator. The fully 
processed molten salt, on its way back to the primary loop, has uranium added 
at the rate required to maintain or adjust the fissile material concentration 
and, hence, the reactivity, in the reactor core as desired.

@@

Figure~\ref{fig:spectrum} represents the normalized neutron flux distribution 
of the initial and the equilibrium core compositions. The spectrum for the 
equilibrium state is harder than the initial state due to heavy fission 
products.  \begin{figure}[hb] % replace 't' with 'b' to force it to be on the 
        bottom
        \centering
        \vspace{-0.85em}
        \includegraphics[width=1.04\linewidth]{spectrum.png}
        \vspace{-0.6em}
        \caption{Normalized neutron energy spectrum for initial and equilibrium 
        composition.}
        \label{fig:spectrum}
        \vspace{-0.10em}
\end{figure}

>>>>

The thorium-fueled \gls{MSBR} was developed in the early 1970s by \gls{ORNL} 
specifically to realize the promise of the thorium fuel cycle which allows the 
use of natural thorium instead of enriched uranium as the fertile element. 
Thorium breeds the fissile $^{233}U$ and avoids uranium enrichment 
\cite{robertson_conceptual_1971}. In the matter of nuclear fuel cycle, the 
thorium cycle produces a reduced quantity of plutonium and \glspl{MA} 
compared to the traditional uranium fuel cycle. Consequently, it may 
significantly increase proliferation resistance when the \gls{MSR} operates in the 
breeder regime. The \glspl{MSR} also could be employed as a converter reactor for 
transmutation or spent fuel from current \glspl{LWR}.

Recently, interest in \glspl{MSR} is resurgent, with multiple new companies 
pursuing commercialization of \gls{MSR} designs\footnote{Examples include both 
        liquid-fueled molten salt designs
        from Transatomic, Terrapower, Terrestrial, and Thorcon.}
To further develop these \gls{MSR} concepts, strategies for online reprocessing 
and refueling special 
computational analysis methods and codes are needed due to the unique completely different 
physics of liquid-fueled nuclear reactor comparing with traditional, 
solid-fueled, reactors. 
Most of the contemporary nuclear reactor physics codes 
do not able to perform depletion calculations in the online reprocessing 
regime. J.J. Powers (\gls{ORNL}) has suggested a novel method to conducting 
depletion simulation for \gls{MSR} with taking into account the online 
reprocessing and refueling based on the deterministic computer code NEWT in 
SCALE \cite{powers_new_2013}. This approach was later used by Jeong et al. to 
find equilibrium state for \gls{MSBR} and validate it with \gls{MCNP}/CINDER90 
model \cite{jeong_equilibrium_2016}. For the development of \gls{MSBR} 
research, this paper presents the single-cell model developed using a 
continuous-energy the Serpent 2 Monte Carlo reactor physics calculation code 
that was employed to find equilibrium core state, and several calculation 
results including the depletion calculation of the single-cell unit.

@@

\begin{figure}[h] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=0.90\linewidth]{zone_I_mesh.png}
        \caption{MSBR unit cell of Zone I geometry.}
        \label{fig:zoneI}
\end{figure}

@@

\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{keff.png}
        \caption{Infinite multiplication factor during 4 years depletion 
        simulation.}
        \label{fig:keff}
\end{figure}

\FloatBarrier

@@

\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{fuel_composition.png}
        \caption{Normalized number density of major isotopes during 4 years 
        depletion.}
        \label{fig:compos}
\end{figure}

\FloatBarrier

The analysis of the fuel salt composition variation gives clearer information 
about equilibrium core state. Figure~\ref{fig:compos} shows the normalized 
number density of important isotopes that have strong influence on the core 
neutronics at the beginning of each depletion time interval. The number density 
of protactinium is very low (less than $10^{16}$1/cm$^3$) but some small amount 
of it produced during 3-days reprocessing step. It can be seen that the 
multiplication factor reaches the equilibrium state at about 950  days. 
Figure~\ref{fig:rates} represents the rates of materials in online reprocessing 
flows during 4 years of depletion calculations. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\captionsetup[table]{
%  labelsep = newline,
%  name = TABLE, justification=justified,
%  singlelinecheck=false,%%%%%%% a single line is centered by default
%  labelsep=colon,%%%%%%
%  skip = \medskipamount}
%\begin{table}[h!]
%\centering
%\begin{tabular}{p{0.25\linewidth} p{0.3\linewidth} p{0.3\linewidth}} \toprule
%      Isotope & First 400 days     & After 400 days          
%\\ \midrule
%$^{232}$Th  & 3.704$\times10^{12}$ & 3.086$\times10^{12}$
%\\ \midrule
%$^{233}$Pa  & 3.704$\times10^{12}$ & 3.086$\times10^{12}$
%\\ \midrule
%$^{233}$U  & $1.05134\pm0.00116$ & $1.00736$
%\\
%\bottomrule
%\end{tabular}
%  \caption{Materials flows rates during \gls{MSBR} online reprocessing 
%  [1/s*cm$^3$].}
%  \label{tab:rates}
%\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[hb] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{rates_fuel.png}
        \caption{Materials flows rates during online reprocessing.}
        \label{fig:rates}
\end{figure}
From the graph below we can see that to keep reactor critical higher $^{233}$U 
flow rate required for first 400 days of operation and after that much smaller 
uranium-233 flow rate from the protactinium decay tank needed. The $^{232}$Th 
rate slightly decreases over 4 years of operation due to other than $^{233}$U 
fissile materials accumulation.

\begin{figure}[hb] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.03\linewidth]{rates_outflow.png}
        \caption{Materials flows rates for the protactinium decay tank during 
        \gls{MSBR} online reprocessing.}
        \label{fig:outflow}
\end{figure}
As shown in Figure~\ref{fig:outflow}, the tank for protactinium decay 
accumulates $^{233}$Pa approximately 200 days and fresh fuel ($^{233}$U) flow 
established also after 200 days. Uranium produced in the tank by protactinium 
decay is separated by circulation of the salt through a flourinator. The fully 
processed molten salt, on its way back to the primary loop, has uranium added 
at the rate required to maintain or adjust the fissile material concentration 
and, hence, the reactivity, in the reactor core as desired.

@@

Figure~\ref{fig:spectrum} represents the normalized neutron flux distribution 
of the initial and the equilibrium core compositions. The spectrum for the 
equilibrium state is harder than the initial state due to heavy fission 
products.  
\begin{figure}[hb!] % replace 't' with 'b' to force it to be on the bottom
        \centering
        \includegraphics[width=1.04\linewidth]{spectrum.png}
        \caption{Normalized neutron energy spectrum for initial and equilibrium 
        composition.}
        \label{fig:spectrum}
\end{figure}
